\chapter{Consequences and Conclusion}\label{chap:conclusion}

\input{chapters/consequences.tex}

\section{Conclusion}
This report aims to prove \cref{conj:bded_had_pn}, that every $K_t$-minor free graph $G$ can be embedded on $f(t)$ pages. This conjecture remains unsolved. 
As discussed throughout this report, the crucial case to solve are book-embeddings of graphs on projective planes. 

At the start, we have gone through one of the most important theorems in $K_t$-minor free graphs, the Graph Minor Structure Theorem. This report explains briefly the Graph Minor Structure Theorem and the Graph Minor Theorem. 
These tools are used to prove \cref{conj:bded_had_pn},

Next, a thorough discussion of book-embeddings and discussion of treewidth and book-embeddings is given. The report then discusses graphs embedded on surfaces and vortices. A proof of bounding the number of pages when the surface is orientable and the graph contains vortices is stated. This handles almost every ingredient of the Graph Minor Structure Theorem. The crucial case that remains unsolved are the families of graphs embeddable on a nonorientable surface of genus $g$. When $g = 1$, then we can embed these graphs on a bounded number of pages. When $g \geq 2$, no method currently exists to embed these graphs on a book with a bounded number of pages. This report has isolated that nonorientable surfaces of genus $g$ are the crucial case to consider. Therefore, more research must be done on book-embeddings of graphs. 

The main obstruction to proving this conjecture is proving \cref{conj:nonorientable_monochromatic_paths}. If this obstruction can be solved, then we believe that we will have an easier time proving a bound. Currently, no strategies to solving this conjecture have been found and we leave this as an open problem. 
