\chapter{Graph Minors}\label{chap:gmst}
This chapter discusses graph minors and some parameters which bound the size of the largest clique minor in a graph. The parameters that are discussed are treewidth, pathwidth, planar graphs, and graphs on surfaces. 
The Graph Minor Structure Theorem, proven by \textcite{robertsonGraphMinorsXVI2003} is an important characterisation of $K_t$-minor free graphs. The Graph Minor Structure Theorem was developed to solve the Graph Minor Theorem, also by \textcite{robertsonGraphMinorsXX2004}. Many problems of the form: ``Every $K_t$-minor free graph has some bounded parameter, with the bound being $f(t)$'' uses some version of the Graph Minor Structure Theorem in its proof. 


This chapter starts by discussing graph minors. Then each component of the graph minor structure theorem is gone through in detail. Finally, the Graph Minor Theorem, which is what the Graph Minor Structure Theorem was designed to prove, is also discussed.

The Graph Minor Structure Theorem states that every $K_t$-minor-free graph can be decomposed to a set of four ingredients. Individually, each of these components is $K_t$ minor free for some $t$, and combining each of these ingredients forms every $K_t$ minor free graph. 
\begin{itemize}
	\item Graphs on surfaces of fixed genus.
	\item Apex sets.
	\item Clique-sums of graphs.
	\item Vortices.
\end{itemize}

\input{chapters/gmst/minors.tex}

\input{chapters/gmst/widths_book_embeddings.tex}

\input{chapters/gmst/planargraphs.tex}

\input{chapters/gmst/surfaces.tex}

\input{chapters/gmst/gmst.tex}

\input{chapters/gmst/gmt.tex}
\newpage
\input{chapters/gmst/minor_closed_families.tex}