\section{Layout of report}
The structure of the report is as follows.
\begin{itemize}
	\item \cref{chap:gmst} discusses every component of the Graph Minor Structure Theorem and how these components combine to form the Graph Minor Structure Theorem. The components that are discussed in detail are treewidth and graphs on surfaces. Finally, the Graph Minor Theorem is discussed.
	\item \cref{chap:book-embeddings} discusses book-embeddings of graphs of bounded treewidth. There is also a discussion of book-embeddings of graphs with a tree-decomposition with torsos of bounded pagenumber. The papers discussed in this section are by \textcite{hickingbothamStackNumberCliqueSum2023} and by \textcite{ganleyPagenumberTrees2001}. 
	\item \cref{chap:planar} discusses book-embeddings of planar graphs and approaches to add on vortices to the graph. There is a discussion of some methods to add on vortices, but there are no results on embedding vortices on planar graphs. 
	\item \cref{chap:orientable} discusses graphs embedded on orientable surfaces and a book-embedding of graphs on orientable surfaces. \textcite{heathPagenumberGenusGraphs1992} proved that graphs embedded on orientable surfaces of some fixed genus have bounded pagenumber. We show that if some conjectures in \cref{chap:planar} is true, then it implies \cref{conj:bded_had_pn} in the case of orientable surfaces. 
	\item \cref{chap:nonorientable} discusses graphs embedded on non-orientable surfaces. There is a discussion of a proof by \textcite{nakamotoBookEmbeddingProjectiveplanar2015} with embedding projective-planar graphs with a bounded number of pages. There is also a discussion of embedding Klein bottle graphs on a bounded number of pages. Finally, an open problem of embedding graphs on higher non-orientable genus surfaces on a bounded number of pages is discussed. A strengthening of this open problem is also discussed, which implies \cref{conj:bded_had_pn}.
	\item \cref{chap:Future Work} is a brief discussion on future work. We discuss \cref{conj:nonorientable_monochromatic_paths}, which implies \cref{conj:bded_had_pn}. 
\end{itemize}

Readers are expected to have at least an undergraduate understanding in graph theory and point-set topology. 
