% !TEX root = ./thesis.tex

\section{Non-orientable surfaces}

For the nonorientable case, we wish to show that:

\begin{theorem}\label{thm:orientablevortices}
	Let $G$ be a graph where all torsos are $k$ almost-embeddable on a surface of genus at most $g$.  Now suppose there are $p$ vortices which are embedded on $G$. Then $G$ can be embedded on $(2s + 2)(s) + \ell$ pages, where $s = \gpk$. 
\end{theorem}

To prove this for the most general case, the following must be shown:

\begin{conjecture}
	For any graph $G$ with nonorientable genus $g$ on a surface $\Sigma$, $G$ can be embedded on $\leq f(g)$ pages such that for any face $F$ on $\Sigma - G$, the number of monochromatic paths is bounded by some $h(g)$. 
\end{conjecture}

For nonorientable surfaces, this is more complicated. \textcite{heathEmbeddingPlanarGraphs1984} did not sufficiently prove the case for non-orientable graphs. \textcite{nakamotoBookEmbeddingProjectiveplanar2015} showed that every projective-planar graph can be embedded in $9$ pages, which was improved to 6 with a more complex proof. However, we want to prove a book-embedding for all nonorientable surfaces. 

We want to show the following:
\begin{conjecture}\label{conj:embedding_nonorientable_graphs}
	Let $G$ be a graph with a 2-cell embedding on a non-orientable surface $\Sigma$ of genus $g$. Then $G$ can be embedded in $O(g)$ pages.
\end{conjecture}

For the case that $g = 1$, then $G$ can be embedded in 6 pages, from \textcite{ozekiBookEmbeddingGraphs2019}. However, it is not known for higher $g$, and in fact the case when $g = 2$ (Klein-Bottle graphs) was stated to be an open problem. 

We want to show:
\begin{conjecture}\label{conj:nonorientable_monochromatic_paths}
	Every graph embedded on a surface of genus $g$ has a book-embedding of $f(g)$ pages where every face has $h(g)$ monochromatic paths. 
\end{conjecture}
This was shown in the case when the surface is orientable. However, the case when the surface is nonorientable has not been solved yet. This statement is a strengthening of \cref{conj:embedding_nonorientable_graphs}, with an extra condition on the faces of the graph.
This implies the following:

\begin{corollary}
	Every $(g, p, k)$-almost-embeddable graph is embeddable on $4g h(g) (k + 1) + f(g)$ pages.
\end{corollary}

\begin{proof}
	Let $G$ be a $(g, p, k)$-almost-embeddable graph, on a nonorientable surface of genus $g$. By \cref{conj:nonorientable_monochromatic_paths}, there is a book-embedding of $G$ where every face has $h(g)$ monochromatic paths. Then by \cref{lem:vortices_mono_paths}, $G$ can be embedded in $f(g) + h(g) p (k+1)$ pages.
\end{proof}

This implies \cref{conj:bded_had_pn}, in the case when every surface is orientable. 
\begin{corollary}
	Every $\mathcal{G}(g, p, k, a)$ graph can be embedded on $2s^2 + 4s + 1$ pages, where $s =  4g h(g) (k + 1) + f(g) + a$. 
\end{corollary}

\begin{proof}
	Let $G \in \mathcal{G}(g, p, k, a)$. Then $G$ has a tree-decomposition where every torso is $(g, p, k, a)$-almost embeddable. From \cref{thm:apex_vertices_pagenumber}, every torso is embeddable on $s = 4g h(g) (k + 1) + f(g) + a$ pages. Then from \cref{corr:bded_pn_tree_decomp}, $G$ is embeddable on $2s^2 + 4s + 1$ pages.
\end{proof}