
\section{Graph Minor Structure Theorem}\label{sec:Kt_Minor_Free}
What is the structure of $K_t$-minor free graphs? We shall show that we can roughly characterise all $K_t$-minor free graphs as graphs that are products of a series of operations. This classification comes from \cite{robertsonGraphMinorsXVI2003}.
\subsection{Clique-minor-free minor-closed families}\label{ssec:Kt_Minor_Closed_families}
We define $\had(G)$ to be the largest $t$ such that $G$ has a $K_t$ minor. 
\subsubsection{Planar graphs}\label{sssec:K_5-free_Planar}
\begin{theorem}\label{thm:K5_Free_Planar}
	If $G$ is a planar graph, then $G$ is $K_5$-minor-free.
\end{theorem}
\begin{proof}
	If $G$ is planar with $n$ vertices and $m$ edges, then we have that $m \leq 3n -6$. However, we have that $K_5$ has $5$ vertices and $10$ edges, but we have that $ 10 > 3 \times 5 - 6$, so $K_5$ is not planar. As the family of planar graphs is minor-closed, then if $G$ is planar, then $K_5$ is minor-free.
\end{proof}

\subsubsection{A short discussion of topology}\label{sssec:topology}
This section comes from Mohar and Thomassen's book on Graphs on Surfaces. \cite{moharGraphsSurfaces2001} A surface $\Sigma$ is a topological space which, at every point, has a neighbourhood homeomorphic to a disk. There are three important surfaces to know- the sphere $S^2$, the torus $T$, and the real projective plane $P$.
\par
We \textit{add} a \textit{handle} to a surface $\Sigma$ by removing two disks in $\Sigma$ and identifying the boundaries such that one goes clockwise and the other goes counterclockwise. We add a \textit{crosscap} by removing a disk in $\Sigma$ and identifying opposite points on the boundary. We add a \textit{twisted handle} to a surface $\Sigma$ by removing two disks in $\Sigma$ and identifying the boundaries such that both go clockwise.
\par
\begin{definition}[Euler Genus]
	The \textit{Euler genus} of a surface $\Sigma$ obtained from a sphere by adding $h$ handles, $c$ crosscaps and $t$ twisted handles is $2h + 2t + c$.
\end{definition}

\begin{example}
	Here are the Euler genus of some important surfaces.
	\begin{enumerate}
		\item The Euler genus of the sphere is $0$.
		\item The Euler genus of the torus is $2$.
		\item The Euler genus of the projective plane is $1$. 
		\item The Euler genus of Klein bottles is $2$. 
	\end{enumerate}
\end{example}

Note that ``genus'' and ``Euler genus'' are two distinct concepts in topology. In this paper, when we discuss genus, we will always discuss \underline{Euler genus}.

We say a surface $\Sigma$ is \textit{orientable} if $\Sigma$ can be obtained from $S^2$ by only adding handles. An example of an orientable surface is the torus.

We say a surface $\Sigma$ is \textit{non-orientable} if $\Sigma$ can be obtained from $S^2$ by adding at least one crosscap or twisted handle. An example of a non-orientable surface is the projective plane or the Klein bottle. 

\subsubsection{Genus-g graphs}\label{sssec:Graph_genus}

We define the \textit{Euler Genus} of a \textit{graph} $G$ to be the smallest Euler genus $g$ surface $\Sigma$ such that $G$ can be embedded on $\Sigma$ without crossings (note that $\Sigma$ can be orientable or nonorientable).

Let $|F(G)|$ be the number of faces in a graph $G$. Then we have that $|V(G)| - |E(G)| + |F(G)| = \chi = 2 - g$. 

\begin{theorem}[Bounded genus]\label{thm:bounded_genus_kt_free}
	If $G$ is a genus $g$ graph, then $G$ is $K_t$-minor free, where $t > \sqrt{6g} + 4$. 
\end{theorem}
This proof mimics the above proof for planarity, but with higher dimensions. 
We can show that if $G$ has genus $g$, then if $G$ has $n$ vertices and $m$ edges, then $n - m + f = \chi = 2-g$, then as each face has at most 3 vertices and each edge is incident to two faces, we have that $f \leq 2m/3$. Therefore, $m \leq 3(n + g - 2)$, and if $K_t$ is embeddable on a genus $g$ graph, then $\binom{t}{2} \leq 3 (t + g - 2)$. Thus $t \leq \sqrt{6g} + 4$. So if a graph has genus $g$, then it is $K_t$-minor-free, where $t > \sqrt{6g} + 4$. 

\subsubsection{Apex vertices}\label{sssec:Apex_Vertices}
An apex vertex $a$ is added to a graph $G$ such that it has arbitrary edges. As such, $a$ can \textit{dominate} all other vertices in $G$, meaning that $a$ can be adjacent to all vertices in $G$. 
\begin{theorem}
	If $G$ is $K_t$-minor free, $G$ with the apex vertex $a$ is $K_{t+1}$- minor free. 
\end{theorem}
\begin{proof}
	We shall prove the contrapositive. Suppose $G + a$ has a $K_{t + 1}$ minor. Then $K_{t + 1}$ has a model in $G + a$ and denote the model function as $\rho$. Now let $v$ be the vertex in $K_{t + 1}$ such that $\rho(v)$ contains $a$. Then if we delete $v$ from $K_{t + 1}$ and delete all the vertices in $\rho(v)$ from $G$, then we have that $K_t$ is a minor of $G'$, where $G'$ is $G - \rho(v)$. But $G'$ is a minor of $G$, as $G'$ does not contain $a$. But this means that $G$ has a $K_t$ minor. 
\end{proof}
\subsubsection{Clique-sums}\label{sssec:Clique_Sums}
The \textit{$k$-clique-sum} of two graphs $G$ and $H$, denoted as $G \# H$, is the graph obtained by performing a series of operation on the cliques of $G$ and $H$. We find cliques in $G$ and $H$, $C_G$ and $C_H$ respectively, such that $C_G$ and $C_H$ have size $k$. Then we identify the vertices in $C_G$ and $C_H$ to glue $G$ and $H$ together. Call this new clique $C$. We can delete some edges between vertices in $C$. \todo{Add picture}


\begin{lemma}
	If $G = G_1 \# G_2$,then $\had(G) = \max(\had(G_1), \had(G_2))$ and $\tw(G) = \max(\tw(G_1), \tw(G_2))$.
\end{lemma}

\begin{example}\label{ex:clique_sum_genus}
	If $G$ is the clique-sum of Euler genus $g$ graphs, then $G$ is $K_{\sqrt{6g} + 5}$-minor-free, but $G$ possibly has unbounded genus.
\end{example}

\begin{theorem}[Wagner's theorem\cite{wagnerUeberEigenschaftEbenen1937}]\label{thm:WagnersTheorem}
	If $G$ is $K_5$-minor-free, then $G$ can be obtained from $(\leq 3)$-clique-sums of planar graphs and the Wagner graph $V_8$.
\end{theorem}

The Wagner graph is given below.
\tikz
	\graph [nodees = {draw, circle}, clockwise, n = 8] {
	subgraph I_n
	}


\subsubsection{Torsos and adhesion}\label{sssec:Torsos and Adhesion}
Given a graph $G$ and a tree-decomposition $\tree$, the \textit{torso} of a bag $B_x$ of $T$ is the graph $G\langle B_x \rangle$, obtained from $G[B_x]$ where $vw$ is a vertex in $G\langle B_x \rangle$ iff $v,w \in B_x \cap B_y$, where $y$ is any neighbour of $x$ in $T$. So the set $B_x \cap B_y$ for all neighbours $y$ of $x$ in $T$ is a clique in $G\langle B_x \rangle$.
We refer to the \textit{adhesion set} as the set $B_x \cap B_y$. 
The \textit{adhesion} of a tree is defined as $\max(|B_x \cap B_y|)$ where $xy$ is an edge in $T$.

\subsubsection{Relationship between clique-sums and treewidth}
Given $G$ and a tree-decomposition $\tree$, we can say that $G$ is a clique-sum of the torsos $G\langle B_x \rangle$ where the size of the cliques that we paste over is at most the adhesion.
\todo{Expand on this!}

\subsubsection{Vortices}\label{sssec:vortices}
Let $G$ be embedded on a surface $\Sigma$, and let $F$ be a face on $G$. Let $D$ be a disc in $\Sigma$ such that $D$ only intersects $G$ only on vertices on the boundary of $F$. We denote these discs as $G$-clean. 

Then let $\Lambda = (x_1, x_2, ..., x_b)$ be a tuple of vertices on the boundary of $F$ such that they intersect $D$. Then we define a new graph $H$ such that $V(G) \cap V(H) = \Lambda$, and there is a \textit{path-decomposition} of $H$ of bags $B_1, B_2, ... B_b$ such that $x_i \in B_i$ for all $i$. $H$ is denoted as a \textit{$D$-vortex} of $G$. The width of a $D$-vortex is the width of the path above, or $\max_i(|B_i| - 1)$. 

\todo{Draw graph with large treewidth, torso and adhesion, but with bounded $K_t$.}
\subsection{Robertson-Seymour Graph Minor Structure Theorem\cite{robertsonGraphMinorsXVI2003}}\label{ssec:Robertson_Seymour_Graph_Structure}
Given $g, p, a \geq 0$, $k \geq 1$, a graph $G$ is $(g, p, k, a)$- almost embeddable if there exists an $A \subseteq V(G)$ with $|A| \leq a$, and there exists subgraphs $G_0, G_1, ...,  G_{p'}$ of $G$ such that:
\begin{itemize}
	\item $G - A = G_0 \cup G_1 \cup G_2 ... G_{p'}$
	\item $p' \leq p$
	\item There is an embedding of $G_0$ onto a surface $\Sigma$ of genus $\leq g$
	\item There exists pairwise disjoint $G_0$-clean discs $D_1, D_2, ..., D_{p'}$ in $\Sigma$
	\item $G_i$ is a $D_i$-vortex of width at most $k$.
\end{itemize}

\begin{theorem}[Robertson-Seymour Graph Minor Structure Theorem for $K_t$-minor-free graphs]
	For all $t$, there exists $g, p, a \geq 0$, $k, \ell \geq 1$, such that every $K_t$-minor-free graph has a tree-decomposition of adhesion $\leq \ell$ and each torso is $(g, p, k, a)$-embeddable. We refer to the family of graphs which satisfy these constants as $\mathcal{G}(g, p, k, a)$. 
\end{theorem}
In fact, there exists a function $t(g, p, k, a)$ such that if a graph has a tree-decomposition of adhesion $\leq \ell$ and each torso is $(g, p, k, a)$-almost embeddable, then $G$ has no $K_t$ minor. Joret and Wood\cite{joretCompleteGraphMinors2013} found that
\begin{theorem}[Bounds on Graph Minor Structure Theorem\cite{joretCompleteGraphMinors2013}]\label{thm:graph_structure_bound_theorem}
	For all graphs $G$,
	$\had(G) \leq a + 48(k + 1)\sqrt{g + p} + \sqrt{6g} + 5$. Moreover, there exists an integer $n \geq a + 1 4 k\sqrt{p + g}$ such that $K_n$ is a minor of some $\mathcal{G}(g, p, k, a)$ graph.
\end{theorem}

\section{Graph Minor Theorem}\label{sec:Graph Minor Theorem}
We move on to the most important and deepest theorem in structural graph theory, the Graph Minor Theorem. This proof was proven in a series of 20 papers by Robertson and Seymour, from 1983 to 2004. As part of the proof, the Graph Minor Structure Theorem was developed, as we have outlined above. 
\begin{theorem}[Graph Minor Theorem \cite{robertsonGraphMinorsXX2004}]
	Every infinite family of graphs contains two distinct graphs $G$ and $H$ such that $H$ is a minor of $G$.
\end{theorem}
Equivalently, this theorem states that:
\begin{theorem}
	Every minor-closed family of graphs can be characterised by a finite set of forbidden minors.
\end{theorem}
Importantly, graphs of bounded genus can be characterised as a set of forbidden minors.
For planar graphs, the two forbidden minors are $K_5$ and $K_{3,3}$. This was proven by Wagner \cite{wagnerUeberEigenschaftEbenen1937}. 
For graphs that can be embedded on a torus, there are at least 17,523 graphs which are forbidden minors, with a database maintained by Myrvold and Woodcock\cite{myrvoldLargeSetTorus2018}. 