\section{Bounds of pagenumbers of graphs}\label{sec:BoundedPagenumber} 
\subsection{Tree-decomposition into bounded page number components}\label{ssec:Clique_sum_Pagenumber_bound}

This proof has been adapted into the language of tree-decompositions. 
\begin{theorem}[Hickingbotham and Wood \cite{hickingbothamStackNumberCliqueSum2023}]\label{thm:clique_sum_pagenumber_bound}
	Let $G$ be a graph with a tree-decomposition $(B_x: x \in V(T))$ where each torso $G \langle B_x \rangle$ has pagenumber $\leq s$ and every torso $G \langle B_x \rangle$ is $c$-colourable. Additionally, we have that the adhesion of this tree is at most $\ell$.
	Then $\pn(G(\mathcal{G}, T)) \leq cs + \ell$.  
\end{theorem}

\subsubsection{Proof of above theorem.}
This proof will involve gluing torsos along adhesion sets, which are cliques. However, the size of these cliques is at most $\ell$. 

Let $C$ be a clique in $G$ and let $\sigma_C = (u_1, ... , u_k)$ be a vertex ordering of $V(C)$, and let $C \leq \ell$. For any arbitrary clique $J$, we define a rainbow-vertex $w \in V(J)$ as a vertex where for any $x, y \in V(J)$, the edges $wx$ and $wy$ are on different pages. We want the book embedding to have the structure $(\underbrace{u_1, u_2, ..., u_k}_{\text{Vertices in } C}, \underbrace{v_1, v_2, ..., v_l}_{\text{Vertices not in }C})$. 

To prove this theorem, we will use a common technique in graph theory. We will strengthen the lemma so that we may use induction to prove the statement.
\begin{lemma}\label{lem:Hickingbotham_Lemma}
	Let $G$ be a graph where $\pn(G) \leq s$ and $\chi(G) \leq c$, and a clique $C$ with an ordering $\sigma_C$. Let $|C| \leq \ell$. There exists a $cs + \ell$-stack layout $(\leq, \psi)$ of $G$ where:
	\begin{enumerate}
		\item The vertex ordering has the structure $(\underbrace{u_1, u_2, ..., u_k}_{\text{Vertices in } C}, \underbrace{v_1, v_2, ..., v_l}_{\text{Vertices not in }C})$. 
		\item For every $u \in V(C)$, the edges $\lbrace uv \in E(G) : u \leq v \rbrace$ are a monochromatic star. 
		\item For every clique $J$, the last vertex of $J$ is a rainbow-vertex. 
	\end{enumerate}
\end{lemma}
\begin{proof}[Proof of \cref{lem:Hickingbotham_Lemma}]
	Let $(\leq_a, \psi_a)$ be a $s$-stack layout of $G$ and let $\rho: V(G) \rightarrow \lbrace 1, 2, ..., c\rbrace$ be a proper colouring of $V(G)$.
	
	Let $u_1, ..., u_k$ be the vertices of $C$ ordered by $\sigma_C$. Note that $k \leq \ell$. Then the new ordering starts with $u_1 \leq u_2 \leq ..., \leq u_k$, and all vertices not in $K$ are placed after, according to $\leq_a$.
	Then the stack assignment $\psi$ is now defined. For every edge $u_i v$ where $u_i \in V(C)$ and $u_i \leq v$, define $u_i v = i$. Otherwise, if neither $u$ or $v$ are in $V(C)$, and $u \leq v$, then let $\psi(uv) = (\rho(u), \psi_a(uv))$. Then we have at most $|\rho| |\psi_a| + k \leq cs + \ell$ pages.
	
	We shall show that $(\leq, \psi)$ is a proper book-embedding. Suppose we have a pair of edges $uv$ and $xy$ which cross, and $\phi(uv) = \phi(xy)$. Suppose that $u$ is the smallest vertex in the ordering $\leq$. If $u \in V(C)$, then the edge $uv$ is assigned to its own page, meaning that it cannot cross $xy$. So $x = u$, but we can draw $uv$ and $uy$ on a single page. Thus they do not cross. Therefore we have that $u, v, x, y$ are not in $V(C)$, and as we preserve the original book-embedding, then these edges do not cross. Thus shown.
	We have that property 1 and 2 are immediate. For property 3, consider a clique $J$ in $G$. Then we must show the last vertex of $J$ is rainbow. Suppose the last vertex of $J$ is $w$, and let $u, v$ be earlier vertices. Since $u$ and $v$ necessarily are assigned different colours in the colouring, then $\psi(uw) = (\rho(u), \psi_a(uw))$ and $\psi(vw) = (\rho(v), \psi_a(vw))$. Therefore, the two edges are on different pages. Thus $w$ is a rainbow vertex. 
\end{proof}

\subsubsection{Full proof}
\begin{theorem}
	Suppose $G$ has a tree-decomposition $(B_x: x \in V(T))$ with torsos $G \langle B_x \rangle$ and adhesion at most $\ell$. Order the vertices of $T$ with a breath-first search, and relabel the vertices $v_0, ..., v_k$ with respect to the bfs ordering. Define $G_i := G \langle B_{v_i} \rangle $. Suppose that for all torsos $G_i$, $i \in [k]$, we have that $\pn(G_i) \leq s$ and $\chi(G_i) \leq c$. Then there is a book-embedding of $G$ with pagenumber of at most $cs + \ell$. 
\end{theorem}
Recall that for a breadth-first search, we maintain the property that for all $i$, $T[v_0, ..., v_{i - 1}]$ is also a tree and $v_i$ is adjacent to one of $v_0, ..., v_{i}$. 
\begin{proof}
	To prove the statement, we shall prove the stronger statement that there exists a book-embedding with the properties described with the lemma above using induction. In particular, we will have that the last vertex of any clique $J$ is a rainbow vertex.
	
	Suppose $k = 0$. Then $G_0$ is a single graph with $\pn(G) \leq s$. Choose $C$ to be an arbitrary vertex $v$ in $G_0$. Then by the lemma above, there is a book-embedding with pagenumber at most $cs + 1$ and every last vertex in a clique is a rainbow vertex.
	
	Suppose $k = n$. Let $C$ be the adhesion clique between $G_n$ and the rest of $G$, where $G_n$ is a leaf of the tree $T$. Denote the induced subgraph $G[V(G) - V(G_n - C)]$ as $G'$. Then let $(\leq_n, \psi_n)$ be the $cs + \ell$-page-number book-embedding of $G_n$ that starts with $V(C)$, and let $\sigma_C$ be the ordering of $V(C)$ by $\leq_n$. Let $(\leq_{n-1}, \psi_{n-1})$ be the stack-embedding of $G'$. By induction, this has a $cs + \ell$-page book-embedding with the properties described above.
	
	\paragraph{Construction of new book-embedding}
	Let $w \in V(K)$ be the last vertex of $K$ with respect to $\leq_{n-1}$. Then insert $V(G_n - C)$ between $w$ and its successor in the order of $\leq_{n}$. For the page assignment $\psi$, we have that if $uv \in E(G')$, then $\psi(uv) = \psi_{n-1}(uv)$. For the remaining edges, we can permute the edge assignments of $\psi_n$ such that for all $u \in V(K)$, we have that $\psi(E_u) = \psi_n(uw)$. We can do this as $w$ is a rainbow vertex and the edges $E_u$ are assigned to a unique page in $\psi_n$. Finally, let $\psi(uv) = \psi_n(uv)$ for the remainder of the edges. Denote the new ordering and assignment as $(\leq, \psi)$. 
	\paragraph{Proof that this is a valid book-embedding}
	We claim that $(\leq , \psi)$ is a stack layout that satisfies the induction hypothesis. Suppose that $\psi(uv) = \psi(xy)$. If $uv, xy \in E(G')$, then by the induction hypothesis, they do not cross. Similarly, if $uv, xy \in E(G_n)$, then they will not cross, by the above lemma. If $uv$ is in $E(G')$ and $xy \in E(G_n)$, then they will go over each other or be sequential and therefore will not cross. 
	Finally, if $u, v, x, y \in C$, then by the induction hypothesis on $G'$, they do not cross either. The final case is if $uv \in E(G_{i + 1})$ and $u \in V(C)$, $v \notin V(C)$, $xy \in E(G')$. If $uv$ and $xy$ cross, then we have that $xy$ and $uw$ will cross. But this will contradict the page-embedding of $G'$.
	
	Let $J$ be a clique in $G$, and $w$ is its final vertex. Then if $J$ is in $G'$, then $w$ is a rainbow-vertex. Otherwise, the last vertex is contained in $G_n$. By the above lemma, $w$ must also be a rainbow vertex. Thus shown.
\end{proof}

\subsubsection{Bounds on pagenumber of $\ell$ and $c$}

We have some bounds in terms of pagenumber on some of these constants, however these constants are not always tight. In particular, we can get better bounds for planar graphs. 

\begin{lemma}[Bounded pagenumber implies bounded clique-number]
	If $pn(G) \leq s$, then $G$ does not contain any cliques on more than $2s+1$ vertices.
\end{lemma}

\begin{proof}
	If $G$ has a clique of size $k$, then embedding the clique of size $k$ and therefore $G$ requires at least $\lfloor \frac{k}{2} \rfloor$ pages, from \cref{thm:Pagenumber_Complete_Graph}. Therefore, if we can embed $G$ in $s$ pages, then the largest clique of $G$ is at most $2s + 1$. Therefore, $\ell \leq 2s + 1$. 
\end{proof}
As we have the bound of $\chi(G) \leq 2 \pn(G) + 2$, from \cref{thm:Colouring_Bound}, we have a bound that does not depend on the chromatic number of $G$. 
\begin{corollary}[Bounded pagenumber of tree-decompositions]
	Let $G$ be a graph with a tree-decomposition $(B_x: x \in V(T))$ where each torso $G \langle B_x \rangle$ has pagenumber $\leq s$. Then from \cref{thm:clique_sum_pagenumber_bound}, $G$ has pagenumber at most $2s^2 + 4s + 1$. 
\end{corollary}

\subsubsection{Bounds on pagenumbers of planar graphs}
This theorem also tells us that pagenumbers of planar graphs are bounded.

\begin{theorem}\label{thm:Planar Graph Hickingbotham Bound}
	Let $G$ be a planar graph. Then $\pn(G) \leq 11$.
\end{theorem}

We will use the corollary \cref{corr:planar_graphs_4_connected_cliqesums}. We will also use the fact that all planar graphs are $4$-colourable \cite{appelEveryPlanarMap1989} and the fact that all 4-connected planar graphs are Hamiltonian, from Tutte \cite{tutteTheoremPlanarGraphs1956}.

\begin{theorem}[Tutte\cite{tutteTheoremPlanarGraphs1956}]\label{thm:4-connected_planar_ham_cycle}
	All 4-connected planar graphs are Hamiltonian.
\end{theorem}

\begin{lemma}\label{lem:clique_sum_connected}
	All graphs are clique-sum trees of $k$-connected subgraphs with adhesion at most $k-1$.
\end{lemma}
\begin{proof}
	We have that from Menger's theorem that if a graph $G$ is not $k$-connected, we can find a separator of size at most $k-1$ such that we have sets $A$ and $B$. Then we repeat this operation on $A$ and $B$ and we can construct a clique-sum tree where every component is $k$-connected and the adhesion size is at most $k-1$. 
\end{proof}

\begin{corollary}\label{corr:planar_graphs_4_connected_cliqesums}
	All planar graphs $G$ have tree-decompositions with the torsos being $4$-connected planar graphs and adhesion at most $3$.
\end{corollary}


\begin{proof}[Proof \cref{thm:Planar Graph Hickingbotham Bound}]
	Let $G$ be planar. Then $G$ has a tree-decomposition of adhesion size at most $3$ with the torsos being $4$-connected, from \cref{lem:clique_sum_connected} However, this implies that the torsos are Hamiltonian, from Tutte \cite{tutteTheoremPlanarGraphs1956}, thus the number of pages needed for each torso is $2$. Therefore from \cref{thm:clique_sum_pagenumber_bound}, we have that the pagenumber is at most $2 \times 4 + 3 = 11$. 
\end{proof}

We will discuss the $K_5$-minor free case. If $G$ is $K_5$-minor free, then we do not need to worry about the GMST to form an upper bound. We will use Wagner's theorem.

\begin{theorem}[Wagner's theorem \cite{wagnerUeberEigenschaftEbenen1937}]
	Every $K_5$-minor free graph is the clique sum of planar graphs and Wagner graphs, where the size of the cliques we sum over is at most 3.
\end{theorem}
Equivalently, if $G$ is $K_5$ minor free, then $G$ has a tree-decomposition $\tree$ where $\tree$ has adhesion at most 3, and every torso of $\tree$ is either planar or the Wagner graph. We will use this to prove the toerem below. 
\begin{theorem}
	If $G$ is $K_5$-minor free, then $G$ has pagenumber $\leq 19$.
\end{theorem}

\begin{proof}
	Suppose $G$ is $K_5$-minor free. Then by Wagner's theorem \cite{wagnerUeberEigenschaftEbenen1937}, $G$ has a tree-decomposition of adhesion at most 3 where every torso is either a planar graph or the Wagner graph.
	We have that planar graphs are $4$-colourable and have pagenumber $\leq 4$. We have that the Wagner graph is $3$-colourable and has pagenumber $\leq 4$. Therefore, we have that if $G$ is $K_5$-minor free, then $G$ has pagenumber at most $4 \times 4 + 3 = 19$. 
\end{proof}
We refer to \cref{fig:wagner} for a description of a circular ordering of a Wagner graph and its colouring. 