% !TEX root = ./thesis.tex
\chapter{Introduction}\label{sec:introduction}
Structural graph theory is a fundamental topic in graph theory. Many results from structural graph theory decompose graphs, or families of graphs, into smaller graphs with bounded parameters. One of the most important theorems in structural graph theory is \textcite{robertsonGraphMinorsXX2004} Graph Minor Theorem which states that every proper minor-closed graph family is characterised by a finite set of forbidden minors. Furthermore, structural graph theory is a field of interest, with connections to topological graph theory, extremal graph theory and algorithmic complexity. 

Topological graph theory is also a fundamental topic in graph theory. The main questions topological graph theory aims to solve are graphs embedded on topological spaces. Topological graph theory is used in data science, optimisation, and computer science. The topological spaces that this report is concerned on are surfaces and quotient spaces of surfaces. 

\input{chapters/intro/problemstatement.tex}

\input{chapters/intro/problemsupport.tex}
\input{chapters/intro/layout.tex}
\newpage
\section{Background}\label{sec:background}
This section is a brief overview of some concepts in topological graph theory and structural graph theory. The concepts discussed are planar graphs, graphs embedded on surfaces, graphs on books, graph minors and the Graph Minor Structure Theorem. The Four Colour Theorem and the Map Colour Theorem are also discussed. These were conjectures for a very long time before being solved in the 1970s. The attempts to prove these theorems helped motivate the study of graphs on surfaces.
\input{chapters/intro/Basic definitions.tex}

\input{chapters/intro/planargraphs.tex}

\input{chapters/intro/surfaces.tex}

\input{chapters/intro/bookembeddings.tex}

