\subsection{Bounding the number of pages of a planar graph}
This subsection uses \cref{lem:planar_graphs_4_connected_cliqesums} and \cref{thm:clique_sum_pagenumber_bound} to find a book-embedding of 4-connected planar graphs. This book-embedding is different from \textcite{yannakakisEmbeddingPlanarGraphs1989} as it does not require a planar triangulation. Because of this fact, this proof is used in future sections with respect to adding vortices on faces. 

Firstly, there is a theorem of Tutte which can be used to bound the number of pages of a $4$-connected planar graph.

\begin{theorem}[Tutte\cite{tutteTheoremPlanarGraphs1956}]\label{thm:4-connected_planar_ham_cycle}
	Every 4-connected planar graph is Hamiltonian.
\end{theorem}

\cref{thm:4-connected_planar_ham_cycle} is used to prove \cref{thm:Planar Graph Hickingbotham Bound}.

\begin{corollary}\label{thm:Planar Graph Hickingbotham Bound}
	Let \(G\) be a 2-connected planar graph. Then $G$ can be embedded on $11$ pages, with book-embedding $(<, \rho)$. $<$ restricted to the outer cycle $C$ is $C$. Furthermore, for every face cycle $C$, $<_{V(C) - \{u, v, w\}} = C - \{u, v, w\}$ for some vertices $u$, $v$, $w$. 
\end{corollary}
\begin{proof}
	From \cref{thm:4-connected_planar_ham_cycle}, every $4$-connected planar graph is Hamiltonian. Furthermore, $K_4$ is a Hamiltonian planar graph.
	Recall that Hamiltonian planar graph can be embedded on two pages, from \cref{lem:Pagenumber_2}. 
	Then apply \cref{thm:clique_sum_pagenumber_bound} with tree-decomposition from \cref{lem:planar_graphs_4_connected_cliqesums} to $G$. Then $G$ can be embedded on \(2 \cdot 4 + 3 = 11\) pages. This has the property that the embedding restricted a face boundary preserves the ordering of the facial walk.

	Furthermore, from the construction given in \cref{lem:planar_graphs_4_connected_cliqesums}, every $4$-connected component is is glued to another $4$-connected component by a separating face. Therefore, every face only changes by $3$ vertices, from \cref{thm:clique_sum_pagenumber_bound}. Therefore removing $3$ vertices from every face preserves the cyclic ordering of every face.
\end{proof}

