\documentclass[]{report}
\usepackage[margin = 1in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[english]{babel}
\usepackage{url}
\usepackage{todonotes}
\usepackage{csquotes}

\usepackage{hyperref}
\usepackage[noabbrev, capitalise]{cleveref}

\usepackage{tikz}
\usetikzlibrary{graphs,graphdrawing, graphs.standard}
\usegdlibrary{trees}
\usepackage{svg}
\svgsetup{inkscapeexe="C:/Program Files/Inkscape/bin/inkscape.exe"}

\usepackage{graphicx}

\usepackage[style = numeric,
isbn=false,
url=false,
eprint = false,
maxbibnames=99
]{biblatex}
\renewbibmacro{in:}{}
\DeclareSourcemap{
	\maps[datatype=bibtex]{
		\map{
			\step[fieldset=url, null]
			\step[fieldset=extra, null]
			\step[fieldset=urldate, null]
			\step[fieldset=location, null]
		}
	}
}
\AtEveryBibitem{%
	\clearfield{day}%
	\clearfield{month}%
	\clearfield{endday}%
	\clearfield{endmonth}%
}

\addbibresource{Book-Embeddings.bib}
% Commands
\newcommand{\tree}{\mathcal{T}}
\DeclareMathOperator{\tw}{tw}
\DeclareMathOperator{\had}{had}
\DeclareMathOperator{\pw}{pw}
\DeclareMathOperator{\td}{td}
\DeclareMathOperator{\pn}{pn}
% Environments

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{example}[theorem]{Example}

\numberwithin{theorem}{section}
\numberwithin{equation}{section}

%includeonly
\includeonly{
chapters/introduction,
chapters/definitions,
chapters/knownresults,
	chapters/knownresults/gmst,
	chapters/knownresults/robertsproof,
	chapters/knownresults/heathandistrail,
chapters/proofattempt
}

%opening
\title{Towards a proof that all \(K_t\)-minor-free graphs have bounded pagenumber}
\author{Eric Luu}

\begin{document}

\maketitle
\listoftodos

\chapter{Abstract}\label{abstract}
In this report we outline our current progress in proving that \(K_t\)-minor free graphs have bounded pagenumber. Proving this bound will connect two important concepts in structural graph theory that have been studied extensively for the past 40 years. We will outline the most important theorem in structural graph theory related to \(K_t\)-minor free graphs, the Graph Minor Structure Theorem. We will also introduce some other proofs which will be of use proving this bounded pagenumber conjecture.
\todo{Move basic definitions to start before the main body of introduction}

\setcounter{tocdepth}{1}
\tableofcontents

\include{chapters/introduction}
\include{chapters/definitions}
\include{chapters/knownresults}
\include{chapters/knownresults/gmst}
\include{chapters/knownresults/robertsproof}
\include{chapters/knownresults/heathandistrail}

\include{chapters/proofattempt}

\chapter{Conclusion}\label{chap:conclusion}
We conclude this report by explaining our current progress and our future plans.
The main important result explained is the Graph Minor Structure Theorem~\cite{robertsonGraphMinorsXVI2003} and its application in our particular problem involving \(K_t\)-minor free graphs. We apply the Graph Minor Structure Theorem to the problem and explain some partial results.
We also go through some other important results in \(K_t\)-minor free graphs and pagenumbers, including Heath and Istrail's \cite{heathPagenumberGenusGraphs1992} paper on graphs on surfaces and bounding the pagenumber.
The main result we have attempted to show is the case where graphs are almost embedded on an orientable surface. We can show that if a graph is almost-embeddable on an orientable surface, then the pagenumber of the graph is bounded.
What remains to show is if a graph is almost-embeddable on a non-orientable surface.
We aim to prove this conjecture:
\begin{conjecture}
	Let $G$ be a graph where all torsos are $k$ almost-embeddable on a surface of genus at most $g$. Now suppose there are $p$ vortices which are embedded on $G$. Then $G$ has bounded pagenumber.
\end{conjecture}

The non-orientable analogue of the orientable case is:
\begin{conjecture}
	Let $G$ be a graph embedded on a non-orientable surface of Euler genus $g$. Let $F_1, \ldots, F_p$ be faces on $G$ which are homeomorphic to a disk. Then there exists a book-embedding $(<, \leq)$ such that for all $F_i$, $1 \leq i \leq p$, we have that $F_i$ needs $\leq f(g)$ monochromatic paths.
\end{conjecture}
We aim to replicate the result from \cref{lem:orientablesurfaces_monochromatic_edges} in the non-orientable case and remove the word ``orientable'' from the statement.
This will be much more difficult as we do not have the same tools for orientable surfaces. Heath and Istrail's proof for non-orientable surface page numbers is much different than the orientable case, and we will need to look at that proof in depth and extend their results for our case.

\printbibliography{}
\end{document}
