\documentclass[]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{ytableau}
\usepackage{float}
\ytableausetup{centertableaux}
\allowdisplaybreaks

\newcommand{\tr}{\text{tr}}
\newcommand{\fix}{\text{Fix}}
%opening
\title{Representation Theory- Problem sheet 9}
\author{Eric Luu}

\begin{document}

\maketitle

\section*{Chapter 6}
\subsection*{Question 9}

\subsubsection*{a}
Consider $(4) \vdash 4$. We have that the Young Tableau is $\ydiagram{4}$. 
Consider $\mu = (4)$. We have that the only ribbon tableau is $\begin{ytableau}
	1 & 1 & 1 & 1
\end{ytableau}
$. The height of the $1$-ribbon is 1. 

\begin{equation}
	\chi^{(4)}_{(4)} = (-1)^{1 - 1} = 1
\end{equation}
For $\mu = (3, 1)$
we also have one way to put the ribbons in place, namely
$\begin{ytableau}
	1 & 1 & 1 & 2
\end{ytableau}
$
 We have the product is:
\begin{equation}
	\chi^{(4)}_{(3,1)} = 1
\end{equation}

In fact, we will have only one ribbon-tableau and the tableau only lives in one row, so
\begin{equation}
	\chi^{(4)}_{\mu} = 1
\end{equation}
for all $\mu$.

Consider $(3,1) \vdash 4$. The young tableau is $\ydiagram{3, 1}$. For $\mu = (4)$, the ribbon tableau is $-1$. 
For $\mu = (3,1)$, then we have that there are two tableaus.
\begin{equation}
	\begin{ytableau}
		1& 1 & 1 \\
		2
	\end{ytableau}
	,
	\begin{ytableau}
		1 & 1 & 2\\
		1
	\end{ytableau}
\end{equation}
We have that:
\begin{equation}
	\chi^{(3,1)}_{(3,1)} = 1 \times 1 + (-1) \times 1 = 0
\end{equation}
For $\mu = (2,1,1)$, we have 3 tableaus. 

\begin{equation}
	\begin{ytableau}
		1 & 1 & 2\\
		3
	\end{ytableau}
	,
	\begin{ytableau}
	1 & 1 & 3\\
	2
\end{ytableau}
,
	\begin{ytableau}
	1 & 2 & 3\\
	1
\end{ytableau}
\end{equation}
so 
$\chi^{(3,1)}_{(2,1,1)} = 1 \times 1 \times 1 + 1 \times 1 \times 1 + (-1) \times 1 \times 1 = 1$. 

For $\mu = (2,2)$, we have that the only tableau is:
\begin{equation}
	\begin{ytableau}
	1 & 2 & 2\\
	1
\end{ytableau}
\end{equation}
Therefore, we have that:
$\chi^{(3,1)}_{(2,2)} = 1 \times -1 = -1$. 


Finally for $\mu = (1,1,1,1)$ we have 
\begin{equation}
	\begin{ytableau}
		1 & 2 & 3 \\
		4
	\end{ytableau}
	,
	\begin{ytableau}
	1 & 2 & 4 \\
	3
\end{ytableau}
,
	\begin{ytableau}
	1 & 3 & 4 \\
	2
\end{ytableau}
\end{equation}
Therefore, we have that $\chi^{(3,1)}_{(2,1,1)} = 1 + 1 + 1 = 3$. 


For $\lambda = (2,1,1)$, we have that for $\mu = (4)$, we have that the only tableau is:
\begin{equation}
	\begin{ytableau}
		1 & 1\\
		1\\
		1
	\end{ytableau}
\end{equation}
so $\chi^{(2,1,1)}_{(4)} = 1$.
For $\mu = (3,1)$, we have that the ribbon tableaus are:
\begin{equation}
	\begin{ytableau}
		1 & 1\\
		1\\
		2
	\end{ytableau}
	\begin{ytableau}
		1 & 2\\
		1\\
		1
	\end{ytableau}
\end{equation}
Therefore, we have that $\chi^{(2,1,1)}_{(3,1)} = - 1 + 1 = 0$.

For $\mu = (2,1,1)$, we have that the ribbon tableaus are:

\begin{equation}
	\begin{ytableau}
		1 & 1\\
		2\\
		3
	\end{ytableau}
	,
	\begin{ytableau}
		1 & 2\\
		1\\
		3
	\end{ytableau}
	,
	\begin{ytableau}
		1 & 3\\
		1\\
		2
	\end{ytableau}
\end{equation}
So $\chi^{(2,1,1)}_{(2,1,1)} = 1 - 1 -1 = -1$. 
For $\mu = (2,2)$ the ribbon tableau is:
\begin{equation}
	\begin{ytableau}
		1 & 1\\
		2\\
		2
	\end{ytableau}
\end{equation}
Therefore, $\chi^{(2,1,1)}_{(2,2)} = -1$.

Finally, for $\mu = (1,1,1,1)$, we have that the ribbon tableau is:
\begin{equation}
	\begin{ytableau}
		1 & 2\\
		3\\
		4
	\end{ytableau}
	,
	\begin{ytableau}
	1 & 3\\
	2\\
	4
\end{ytableau}
,
	\begin{ytableau}
	1 & 4\\
	2\\
	3
\end{ytableau}
\end{equation}
Therefore, $\chi^{(2,1,1)}_{(1,1,1,1)} = 1 + 1 + 1 = 3$.


Suppose $\lambda = (2,2)$. We have that $\chi^{(2,2)}_{(4)} = 0$ as there is no ribbon of length $(1,1,1,1)$. 

Then for $\mu = (3,1)$, we have that there are two $\lambda$-tableau which are
\begin{equation}
	\begin{ytableau}
		1 & 1\\
		1 & 2
	\end{ytableau}
	,
	\begin{ytableau}
		1 & 2\\
		1 & 1\\
	\end{ytableau}
\end{equation}
meaning that $\chi^{(2,2)}_{(3,1)} = -1 - 1 = -2$.
For $\mu = (2,1,1)$, we have that the ribbons are:
\begin{equation}
	\begin{ytableau}
		1 & 1\\
		2 & 3
	\end{ytableau}
	,
	\begin{ytableau}
	1 & 2\\
	1 & 3
\end{ytableau}
\end{equation}
and therefore $\chi^{(2,2)}_{(2,1,1)} = 1 - 1 = 0$.

For $\mu = (2,2)$, we have that the ribbons are:
\begin{equation}
	\begin{ytableau}
		1 & 1\\
		2 & 2
	\end{ytableau}
	,
	\begin{ytableau}
		1 & 2\\
		1 & 2\\
	\end{ytableau}
\end{equation}
Therefore $\chi^{(2,2)}_{(2,2)} = 1+ 1 = 2$.
Finally, for $\mu = (1,1,1,1)$ we have that the ribbons are:
\begin{equation}
	\begin{ytableau}
		1 & 2\\
		3 & 4
	\end{ytableau}
	,
	\begin{ytableau}
		1 & 3\\
		2 & 4
	\end{ytableau}
\end{equation}
so $\chi^{(2,2)}_{(1,1,1,1)} = 1+ 1 = 2$.

The sign representation is already known and is trivial to compute for all ribbons.

So $S_4$ character table is:
\begin{table}[h!]
	\centering
	\begin{tabular}{|l|l|l|l|l|l|}
		\hline
		$S_4$    & $(1)$ & $(12)$ & $(123)$ & $(12)(34)$ & $(1234)$ \\ \hline
		$\chi_1$ & 1     & 1      & 1       & 1          & 1        \\ \hline
		$\chi_2$ & 1     & -1     & 1       & 1          & -1       \\ \hline
		$\chi_3$ & 3     & 1      & 0       & -1         & -1       \\ \hline
		$\chi_4$ & 2     & 0      & -1      & 2          & 0        \\ \hline
		$\chi_5$ & 3     & -1     & 0       & -1         & 1        \\ \hline
	\end{tabular}
\end{table}

\subsection*{b}
Have that $\lambda = 2 + 2$. 

\section*{Chapter 7}
\subsection*{Question 1}

\subsubsection*{a}
We have that $S_3(\mathbb{C})$ is the set of $3 \times 3$ matrices with nonzero determinant. 
Then we have that $\mathfrak{sl}_n(\mathbb{C})$ is the tangent space around $I$, or the matrices 
\begin{equation}
	A = 
\begin{bmatrix}
	a & b & c\\
	d & e & f\\
	g & h & i\\
\end{bmatrix}
\end{equation}
 that satisfy:
\begin{equation}
	\det \left(I + \varepsilon A\right) \approx 1.
\end{equation}

Then this implies that:
\begin{equation}
	\det
	\left(
	\begin{bmatrix}
		1 + \varepsilon a & \varepsilon b & \varepsilon c \\
		\varepsilon d & 1 + \varepsilon e & \varepsilon f \\
		\varepsilon g & \varepsilon h     & 1 + \varepsilon i\\
	\end{bmatrix}
	\right)
	= 1
\end{equation}
Expanding this out, we have:
\begin{equation}
	1 + (a + e + i)\varepsilon + O(\varepsilon^2) = 1
\end{equation}
so $a + e + i = 0$. So the tangent space are $3 \times 3$ matrices whose trace is 0.

We have that $SL_n(\mathbb{C})$ is the group of $n \times n$ matrices with nonzero determinant. 

Then we have that $\mathfrak{sl}_n(\mathbb{C})$ is the tangent space around $I$, or the matrices $A$ that satisfy:

\begin{equation}
	\det \left(I + \varepsilon A\right) \approx 1.
\end{equation}
Now we have that the determinant has $\prod_{i = 1}^3 ( 1 + \varepsilon a_{i,i})$ and terms $\prod_{i = 1} \varepsilon A_{\sigma_i}^2$ which disappear. However the terms that disappear are $O(\varepsilon^2)$ and the terms that remain is: $1 + \tr(A) \varepsilon + O(\varepsilon^2) \approx 1$, meaning that we must have that $\tr(A) = 0$. 

Therefore, we have that $\mathfrak{sl}_n(\mathbb{C}) = \left\{A \in \mathbb{C}^{n \times n} | \tr(A) = 0\right\}$.
This holds for $SL_3$ and $\mathfrak{sl}_3$. 

\subsubsection*{b}
Consider $O_2(\mathbb{C})$, the orthogonal group. Then we have that the tangent space of $I$ are matrices $\begin{bmatrix}
	a & b \\
	c & d
\end{bmatrix}
$
that satisfy 
\begin{equation}
	\left(I + \varepsilon \begin{bmatrix}
		a & b \\
		c & d
	\end{bmatrix}\right)^T 
\left(I + \varepsilon \begin{bmatrix}
	a & b \\
	c & d
\end{bmatrix}\right)
 = I
\end{equation}
Then this implies that:
\begin{equation}
	\begin{bmatrix}
		1 + \varepsilon a & \varepsilon c\\
		\varepsilon b & (1 + \varepsilon d)
	\end{bmatrix}
	\begin{bmatrix}
	1 + \varepsilon a & \varepsilon b\\
	\varepsilon c & (1 + \varepsilon d)
\end{bmatrix}
= 
\begin{bmatrix}
	1 & 0\\
	0 & 1
\end{bmatrix}
\end{equation}
However, this means that:
\begin{align*}
	(1 + \varepsilon a)^2 + \varepsilon^2 c^2 &= 1\\
	\varepsilon b + \varepsilon^2 ab + \varepsilon c + \varepsilon^2 cd &= 0\\
	(1 + \varepsilon d)^2 + \varepsilon^2 b &= 1
\end{align*}
Expanding this out, we have that:
\begin{align*}
	1 + \varepsilon 2a + \varepsilon^2 (c^2 + a^2) &= 1\\
	\varepsilon (b + c) + \varepsilon^2 (ab + cd) &= 0\\
	1 + \varepsilon 2d + \varepsilon^2 (b^2 + d^2) &= 1
\end{align*}
meaning that $a = d = 0$ and $b = -c$. 
So $\mathfrak{so}_2(\mathbb{C})$ is the set of $2 \times 2$ skew-symmetric matrices. 

Now let $\mathfrak{so}_n(\mathbb{C})$ on $O(n)\mathbb{C}$. Let $A$ be a matrix in $\mathfrak{so}_n(\mathbb{C})$.
\begin{align*}
	(I + \varepsilon A)(I + \varepsilon A)^T &= I\\
	(I + \varepsilon A)(I + \varepsilon A^T) &= I\\
	I + \varepsilon A^T + \varepsilon A + \varepsilon^2 A A^T &= I\\
	\varepsilon (A + A^T) + O(\varepsilon^2) &= 0
\end{align*}
So this implies that $A = - A^T$, or that $A$ is skew-symmetric. However, we have that $O(n)$ has two connected components topologically, and in fact $SO(n)$ is the component that $I$ lives in. So we call it $\mathfrak{so}_n$ because we only deal with the connected component that $I$ lives in. 



\subsection*{Question 2}
An irreducible representation of a Lie algebra $\mathfrak{gl}(V)$ is one where the only invariant subspace is $\left\{ 0 \right\}$ or $V$. Consider the representation $\mathfrak{sl}_2(\mathbb{C}) \rightarrow \mathfrak{sl}(V_d)$. Now consider the eigenvectors of $E, F, H$. We have that every basis vector is an eigenvector of $H$ but with different eigenvalues, so we cannot take summations to produce a different eigenspace of $H$. But note that $E$ and $F$ have one different eigenvector each, but they send that one to the zero vector. Therefore the invariant subspace of $\mathfrak{sl}_2(\mathbb{C})$ must be either $0$ or $V_d$ itself. Thus $\mathfrak{sl}(V_d)$ is irreducible. 
\subsection*{Question 3}
Let $e_0 = x^d, e_1 = x^{d - 1} y , ... , e_d = y^d$. Then let $E$ be the $d + 1 \times d + 1$ matrix indexed at 0. We have that $E(e_i) = i e_{i - 1}$. Then $F(e_i) = (d - i) e_{i + 1}$. Finally, we have that $H(e_i) = (d - 2i) e_i$. 

Then we have that $E_{i, i - 1} = i$, $F_{i, i + 1} = d - i$ and $H_{i, i} = d - 2i$, and zeroes everywhere else. 

\subsection*{Question 4}
Consider $\phi: \mathfrak{sl}_2(\mathbb{C}) \rightarrow \mathfrak{gl}(V)$, and consider the eigenvectors of $\phi(H)$. We have that $\phi(H) = x \dfrac{d}{dx} - y \dfrac{d}{dy}$ on $V$. Now we look at the eigenvectors of $H$. Now we have that if we can decompose $H$ to $m_0 V_0 \oplus m_1 V_1 \oplus m_2 V_2 \oplus ... $, then the eigenvectors that come out must be the basis vectors of $V_i$ for some $V_i$. In fact, all eigenvectors of $V_i$ of the form $x^{i - j} y^{j}$ have eigenvector $m_i \times (i - 2j)$ for some constant $m_i$. Then $V$ will have $m_i V_i$ in its decomposition. This is how we can decompose $\mathfrak{gl}(V)$. 
\end{document}
