\section{Surfaces and graphs on surfaces}
This section discusses surfaces, and graphs on surfaces. Graphs on surfaces is a natural extension to graphs on planes. Furthermore, graphs on fixed surfaces are an essential ingredient to the Graph Minor Structure Theorem.



\subsection{Surfaces}

This section comes from Mohar and Thomassen's\cite{moharGraphsSurfaces2001} book on graphs on surfaces. A \textit{surface} \(\Sigma\) is a topological space which, at every point, has a neighbourhood homeomorphic to an open disk. There are three important surfaces to know- the sphere \(S^2\), the torus \(T\), and the real projective plane \(P\). 

\textit{Handles} are added to a surface to a surface \(\Sigma\) by removing two disks in \(\Sigma\) and identifying the boundaries such that one goes clockwise and the other goes counterclockwise. We add a \textit{crosscap} by removing a disk in \(\Sigma\) and identifying opposite points on the boundary. 
\par
\begin{definition}[Euler Genus]
	The \textit{Euler genus} of a surface \(\Sigma\), obtained from a sphere by adding \(h\) handles and \(c\) crosscaps, is \(2h + c\).
\end{definition}

\begin{example}
	This is the Euler genus of some surfaces.
	\begin{enumerate}
		\item The Euler genus of the sphere is \(0\).
		\item The Euler genus of the torus is \(2\).
		\item The Euler genus of the projective plane is \(1\). 
		\item The Euler genus of the Klein bottles is \(2\). 
	\end{enumerate}
\end{example}

Note that ``genus'' and ``Euler genus'' are two distinct concepts in topology. In this paper, when we discuss genus, we will always discuss \underline{Euler genus}.

A surface \(\Sigma\) is \textit{orientable} if \(\Sigma\) can be obtained from \(S^2\) by only adding handles. An example of an orientable surface is the torus.

A surface \(\Sigma\) is \textit{non-orientable} if \(\Sigma\) can be obtained from \(S^2\) by adding at least one crosscap or twisted handle. An example of a non-orientable surface is the projective plane or the Klein bottle. 

The \textit{Euler Genus} of a \textit{graph} \(G\) is the smallest Euler genus \(g\) surface \(\Sigma\) such that \(G\) can be embedded on \(\Sigma\) without crossings (note that \(\Sigma\) can be orientable or nonorientable). 


\subsection{Graphs on surfaces}

\begin{theorem}[Euler's formula on surfaces]\label{thm:Euler_surfaces}
	Let \(|F(G)|\) be the number of faces in a graph \(G\). Then \(|V(G)| - |E(G)| + |F(G)| = \chi = 2 - g\). 
\end{theorem}

Given graphs $G$ and $H$ with genus $g_1, g_2$, it is useful to construct a new graph with genus $g_1 + g_2$. 
\begin{theorem}[\textcite{millerAdditivityTheoremGenus1987}]\label{thm:additivity_genus}
	Let graphs $G$ and $H$ have genus $g_1$, $g_2$. Then the graph obtained from identifying a vertex in $G$ to a vertex in $H$ has genus $g_1 + g_2$. 
\end{theorem}

\begin{theorem}\label{thm:K5_Free_Planar}
	If \(G\) is a planar graph, then \(G\) is \(K_5\)-minor-free.
\end{theorem}
\begin{proof}
	If \(G\) is planar with \(n\) vertices and \(m\) edges where $n \geq 3$, then \(m \leq 3n -6\).
	However \(K_5\) has \(5\) vertices and \(10\) edges, but  \( 10 > 3 \times 5 - 6\), so \(K_5\) is not planar. As the family of planar graphs is minor-closed, if \(G\) is planar, then $G$ is \(K_5\)-minor free.
\end{proof}

\begin{theorem}[Bounded genus]\label{thm:bounded_genus_kt_free}
	If \(G\) is a genus \(g\) graph, then \(G\) is \(K_t\)-minor free, where \(t > \sqrt{6g} + 4\). 
\end{theorem}
\begin{proof}
	This proof mimics the above proof for planarity, but on surfaces of higher genus. 
	Suppose \(G\) has \(n\) vertices and \(m\) edges and of genus $g$. Then \(n - m + f = \chi = 2-g\), from \cref{thm:Euler_surfaces}. As at least three vertices bound each face and each edge touches exactly two faces, then \(f \leq 2m/3\). Therefore, \(m \leq 3(n + g - 2)\). If \(K_t\) is embeddable on a genus \(g\) graph, then \(\binom{t}{2} \leq 3 (t + g - 2)\). Thus \(t \leq \sqrt{6g} + 4\). So if a graph has genus \(g\), then it is \(K_t\)-minor-free, where \(t > \sqrt{6g} + 4\). 
\end{proof}